# Reproducible Research: Peer Assessment 1

The following libraries are required

* data.table
* ggplot2
* plyr

Please install them using install.packages function

Also set up your working directory using setwd function.  The following code will download and unzip the file.

```{r echo = TRUE}
if (!file.exists("activity.zip")){
  download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip","activity.zip",method="curl")
}

if (!file.exists("activity.csv")){
  unzip("activity.zip")
}

```



## Loading and preprocessing the data

We load the data using data.table libraries fread function. Secondly i convert the date field from character to data using as.Date function

```{r echo = TRUE}
library(data.table)
data<-fread("activity.csv")
#Convert date column to date field
data$date<-as.Date(data$date,format="%Y-%m-%d")
```



## What is mean total number of steps taken per day?
```{r echo = TRUE}
library(plyr)
total_steps<-ddply(data,"date",summarise,total.steps=sum(steps,na.rm=T))
hist(total_steps$total.steps)
```
### Mean
```{r echo = TRUE}
mean(total_steps$total.steps)
```
### Median
```{r echo = TRUE}
median(total_steps$total.steps)
```
## What is the average daily activity pattern?

### 5 minute interval plot

For 5 minute interval, I use plyr ddply to summarize by interval and calculate the average.

```{r echo = TRUE}
#We need ggplot for this
library(ggplot2)
#Lets summarize the data
interval_steps<-ddply(data,"interval",summarise,avg.steps=mean(steps,na.rm=T))
head(interval_steps)
ggplot(data=interval_steps,aes(x=interval,y=avg.steps))+geom_line()
```

Looking at the plot it looks like most of the activity is during the morning

### The interval with maximum number of steps
It looks like most steps are taken at the 8:35 interval
```{r echo = TRUE}
interval_steps[interval_steps$avg.steps==max(interval_steps$avg.steps),]
```


## Imputing missing values

Total Number of missing values

```{r echo = TRUE}
nrow(data[is.na(steps),])
```

My strategy to impute these missing values is to get the average of the interval and populate the items

```{r echo = TRUE}
#Lets copy our data set in to datax
datax<-data
for (i in 1:nrow(datax)){
  if(is.na(datax[i,steps])){
    datax[i,steps:=as.integer(interval_steps[interval_steps$interval==datax[i,interval],"avg.steps"])]
  }
}
```
### Histogram of Imputed data set
```{r echo = TRUE}
#Lets create a histogram to see if this is different than the one on the top

total_steps_imputed<-ddply(datax,"date",summarise,total.steps=sum(steps,na.rm=T))
hist(total_steps_imputed$total.steps)
```
### Mean of Imputed data set
```{r echo = TRUE}
mean(total_steps_imputed$total.steps)
```
### Median of Imputed data set
```{r echo = TRUE}
median(total_steps_imputed$total.steps)
```

## Are there differences in activity patterns between weekdays and weekends?
### Weekdays
```{r echo = TRUE}
data[["day_of_week"]]<-weekdays(data[,date])
week_days<-data[day_of_week %in% c("Monday","Tuesday","Wednesday","Thursday","Friday")]
wd_interval_steps<-ddply(week_days,"interval",summarise,avg.steps=mean(steps,na.rm=T))
ggplot(data=wd_interval_steps,aes(x=interval,y=avg.steps))+geom_line()
```

### Weekends
```{r echo = TRUE}
week_ends<-data[day_of_week %in% c("Saturday","Sunday")]
we_interval_steps<-ddply(week_ends,"interval",summarise,avg.steps=mean(steps,na.rm=T))
ggplot(data=we_interval_steps,aes(x=interval,y=avg.steps))+geom_line()
```

* By looking at both the plots, its easy to tell that on weekdays the activity is mostly for going to work and coming back, during the day 
there is less activity.

* But on weekends the pattern is activity during most of the day 

* In both cases as we expect the steps are minimal during the night

